\section*{Acknowledgments}

%I would like to thank everyone that supported me, directly or indirectly, through my studies at MIT. 

I would first like to thank my thesis advisor, Professor Asada, who gave me the opportunity to work on unique robotic research projects in the d'Arbeloff laboratory. His guidance was invaluable, always thinking ahead to guide my efforts in a direction that would generate innovative and relevant research. I hope that I have learned some of his wisdom during my years at MIT. Also, special thanks to Professor Hogan and Professor Slotine for been members of my thesis committee: Their advices and feedback contributed to strengthening my thesis. Many thanks to all my labmates at the d'Arbeloff laboratory. The lab was a very dynamic and exciting place to work. Special mention to Lluis and Kosuke for spending precious time helping me reviewing technical drawings of gearbox prototypes. I would also like to thank the whole MIT community, for being a constant source of inspiration. The education I have received during my PhD studies was of outstanding quality, and I learned a lot spending time with members of the MechE and the CSAIL community.

I would also like to thank The Boeing Company, Sumitomo Heavy Industry, \textit{Le Fond Qu\'{e}b\'{e}cois de Recherche sur la Nature et Technologie} (FQRNT) and The Natural Sciences and Engineering Research Council of Canada (NSERC), who contributed to funding my PhD studies at MIT.

I must also acknowledge the strong support I have received from my friends and family. My wife Catherine has given outstanding support, and I could not have accomplished everything I did without her. I must also thank all my friends who came to visit me in Boston; their visits provided me with breaks from the frenetic lifestyle of MIT studies. I would also like to thanks my parent for always encouraging me in my studies, and Jean-S\'{e}bastien for giving me the push I needed to pursue my ambitions. Without them I would never have been able to pursue my dream at MIT. 
